Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-05-29 manuscript.Rmd
Figure and table cross-references

## Analyses
### Cross-Recurrence Quantification Analysis
We first used cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series [@zbilut1992embeddings; @webber1994dynamical], CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system [cf. Nyquist frequency in signal processing; @grenander1959probability]. We converted each time series (e.g., Figure 1) into deciles (e.g., Figure 2) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

#### Visualization
A recurrence plot (RP) is the visual representation of CRQA. To construct a recurrence plot, two time series are plotted against each other on the X and Y axes, and at every point where they share the same state, and point in placed on the plot to indicate recurrence. Visually analyzing these plots can help reveal patterns of structure and recurrence. The line y = x is known as the line of identity. This line falls where the time series match temporally, and a filled line would indicate perfect synchrony between the two time series. See Figures 3-5 for examples of RPs.

@@

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a time series. This allows for visual interpretation of the pattern or fluctuation of the metrics across time. These dynamics can then be linked to other external events. See Figures 9-14 and 16-21 for examples.

@@

### Syria as Target
A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, maxL, and LAM. All other metrics were non-significant (Figure 3; Table 1). A permutation test for the CRQA of social cohesion and count of positive events found no statistical difference from chance for any of the metrics (Figure 4; Table 1). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, ENTR, and TT. All other metrics were non-significant (Figure 5; Table 1). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

@@

# build table
pander(target_results, caption = "CRQA results for target only data.")

@@

### Syria as Source and Target
A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, and LAM. All other metrics were non-significant (Figure 6; Table 2). A permutation test for the CRQA of social cohesion and count of positive events found a statistical difference from chance for DET, NRLINE, maxL, and LAM (Figure 7; Table 2). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, and ENTR. All other metrics were non-significant (Figure 8; Table 2). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

@@

# build table
pander(source_target_results, caption = "CRQA results for source and target data.")

@@

WCRQA was performed using the crqa package [@R-crqa] for R [@R-base]. Permutation tests were conducted for each WCRQA to establish upper and lower 95th and 99th percentiles. In the plots (see Figures 9-14 and 16-21), the 95th percentiles are represented by horizontal red lines and the 99th percentiles are represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there is a consistent period of above chance recurrence rate from the 31st through 48th window (Figure 9). Determinism is also almost consistently above chance from the 39th window onwards (Figure 10). For social cohesion and count of positive events there is a downward trend in RR as the window moves (Figure 11) and determinism stays almost entirely within chance across windows (Figure 12). For social cohesion and count of negative events, recurrence rate stay above or almost above chance from the 36th through the 48th window (Figure 13) and determinism is also above or nearly above chance from the 38th through 43rd window (Figure 14).

@@

For social cohesion and count of all events, there is a consistent period of above or nearly above chance determinism from the 26th through 46th window (Figure 16). Determinism also shows above or nearly above chance values in window spans 11-16 and 35-37 (Figure 17). For social cohesion and count of positive events, there is a downward trend in RR as the window moves (Figure 18) and determinism stays largely within chance with the exception of windows 13-19. For social cohesion and negative events, there appears to be an extended time period where RR (Figure 11) and DET (Figure 12) were greater than chance, from the 35th through 46th windows. These windows span from May 3rd, 2012 to May 14th, 2012.

@@

### Syria as Source and Target
By adding Syria as a source in the data filtering, the dataset increased in size by 1,690 samples. These additional observations are times where Syria initiated an event with another country. Events initiated by Syria but directed internally within the country were already captured by the target only filtering. The increase in coupling between positive events and social cohesion is thus due to the inclusion of these externally directed events. Though CRQA metrics for count of all and count of negative events with social cohesion were largely unaffected, count of positive events and social cohesion showed significant levels of determinism, number of lines, max line length, and laminarity that were not present in the target filtered data. Given that the Syria during this time was still under Authoritarian rule, the government may have been initiating more positive external events to prevent intervention by other countries or try to cover up what was happening internally. We believe that the increased coupling of positive events is thus not reflective of actual social cohesion, but happened by chance. The inflation of the number of positive events when including source observations is causing increased coupling, but it is more likely that the Twitter users were using the platform to raise awareness of the violent, internal events that were being inflicted on them by their government officials and armies and not reporting on the positive external events that did not directly affect them or their safety. Further research should explore the semantic content of this tweet corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs; Figures 3-8), there appears to be a shift in the dynamics of these systems around observation 35, or May 3rd, 2012. To further investigate these local dynamics, we performed a sliding window technique known as windowed cross-recurrence quantification analysis (WCRQA).

## Windowed-Recurrence Quantification Analysis
### Syria as Target
The time period of differing dynamics, from observation 38 through 43, appears to be captured by WCRQA of social cohesion and count of negative events for both RR (Figure 13) and DET (Figure 14). The amount of recurrence and determinism for these periods almost consistently lie above chance, suggesting that the intensity of negative events coincide with the degree of social cohesion found on Twitter, and that these states carry across multiple days in a way that is different from that of chance.
The above-chance levels of determinism and recurrence rate are demonstrating the coupling of social cohesion and counts of all and negative events during times of strife. We can see this string of highly violent events reflected in the mode ICEWS event for each day (Figure 15), where from May 1st, 2020 to May 15th, 2020, the most frequently occurring event was almost consistently -10, the most violent event in the coding system. These dates directly correlate with the above chance RR and DET metrics for the WCRQA of social cohesion and count of negative events. Furthermore, for WCRQA of social cohesion and positive event count during this time period, DET (Figure 12) was entirely within chance. This supports our hypothesis, as the positive events are seemingly not coupled with the social cohesion on Twitter during civil uprisings during the Arab Spring.

>>>>

## Analyses
### Cross-Recurrence Quantification Analysis
We first used cross-recurrence quantification analysis [CRQA; @zbilut1998detecting] to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series [@zbilut1992embeddings; @webber1994dynamical], CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system [cf. Nyquist frequency in signal processing; @grenander1959probability]. We converted each time series (e.g., Figure \@ref(fig:raw-ts)) into deciles (e.g., Figure \@ref(fig:deciled-ts)) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

#### Visualization
A recurrence plot (RP) is the visual representation of CRQA. To construct a recurrence plot, two time series are plotted against each other on the X and Y axes, and at every point where they share the same state, and point in placed on the plot to indicate recurrence. Visually analyzing these plots can help reveal patterns of structure and recurrence. The line y = x is known as the line of identity. This line falls where the time series match temporally, and a filled line would indicate perfect synchrony between the two time series. See Figures \@ref(fig:plot-rp-targ-all)-\@ref(fig:plot-rp-source-targ-neg) for examples of RPs.

@@

#### Visualization
To visualize WCRQA, the resulting metrics from each window can be plotted as a time series. This allows for visual interpretation of the pattern or fluctuation of the metrics across time. These dynamics can then be linked to other external events. See Figures \@ref(fig:plot-wcrqa-targ-all) and \@ref(fig:plot-wcrqa-source-targ-all) for examples.

@@

### Syria as Target
A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, maxL, and LAM. All other metrics were non-significant (Figure \@ref(fig:plot-rp-targ-all); \autoref{table-1}). A permutation test for the CRQA of social cohesion and count of positive events found no statistical difference from chance for any of the metrics (Figure \@ref(fig:plot-rp-targ-pos); \autoref{table-1}). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, ENTR, and TT. All other metrics were non-significant (Figure \@ref(fig:plot-rp-targ-neg); \autoref{table-1}). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

@@

# build table
pander(target_results, caption = "\\label{table-1}CRQA results for target only data.")

@@

### Syria as Source and Target
A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, and LAM. All other metrics were non-significant (Figure \@ref(fig:plot-rp-source-targ-all); \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of positive events found a statistical difference from chance for DET, NRLINE, maxL, and LAM (Figure \@ref(fig:plot-rp-source-targ-pos); \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, and ENTR. All other metrics were non-significant (Figure \@ref(fig:plot-rp-source-targ-neg); \autoref{table-2}). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

@@

# build table
pander(source_target_results, caption = "\\label{table-2}CRQA results for source and target data.")

@@

WCRQA was performed using the crqa package [@R-crqa] for R [@R-base]. Permutation tests were conducted for each WCRQA to establish upper and lower 95th and 99th percentiles. In the plots (see Figures \@ref(fig:plot-wcrqa-targ-all) and \@ref(fig:plot-wcrqa-source-targ-all)), the 95th percentiles are represented by horizontal red lines and the 99th percentiles are represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there is a consistent period of above chance recurrence rate from the 31st through 48th window (Figure \@ref(fig:plot-wcrqa-targ-all)). Determinism is also almost consistently above chance from the 39th window onwards (Figure \@ref(fig:plot-wcrqa-targ-all)). For social cohesion and count of positive events there is a downward trend in RR as the window moves and determinism stays almost entirely within chance across windows (Figure \@ref(fig:plot-wcrqa-targ-all)). For social cohesion and count of negative events, recurrence rate stay above or almost above chance from the 36th through the 48th window and determinism is also above or nearly above chance from the 38th through 43rd window (Figure \@ref(fig:plot-wcrqa-targ-all)).

@@

For social cohesion and count of all events, there is a consistent period of above or nearly above chance determinism from the 26th through 46th window (Figure \@ref(fig:plot-wcrqa-source-targ-all)). Determinism also shows above or nearly above chance values in window spans 11-16 and 35-37 (Figure \@ref(fig:plot-wcrqa-source-targ-all)). For social cohesion and count of positive events, there is a downward trend in RR as the window moves and determinism stays largely within chance with the exception of windows 13-19 (Figure \@ref(fig:plot-wcrqa-source-targ-all)). For social cohesion and negative events, there appears to be an extended time period where RR  and DET were greater than chance, from the 35th through 46th windows (Figure \@ref(fig:plot-wcrqa-source-targ-all)). These windows span from May 3rd, 2012 to May 14th, 2012.

@@

### Syria as Source and Target
By adding Syria as a source in the data filtering, the dataset increased in size by 1,690 samples. These additional observations are times where Syria initiated an event with another country. Events initiated by Syria but directed internally within the country were already captured by the target only filtering. The increase in coupling between positive events and social cohesion is thus due to the inclusion of these externally directed events. Though CRQA metrics for count of all and count of negative events with social cohesion were largely unaffected, count of positive events and social cohesion showed significant levels of determinism, number of lines, max line length, and laminarity that were not present in the target filtered data. Given that the Syria during this time was still under Authoritarian rule, the government may have been initiating more positive external events to prevent intervention by other countries or try to cover up what was happening internally. We believe that the increased coupling of positive events is thus not reflective of actual social cohesion, but happened by chance. The inflation of the number of positive events when including source observations is causing increased coupling, but it is more likely that the Twitter users were using the platform to raise awareness of the violent, internal events that were being inflicted on them by their government officials and armies and not reporting on the positive external events that did not directly affect them or their safety. Further research should explore the semantic content of this tweet corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs; Figures \@ref(fig:plot-rp-targ-all)-\@ref(fig:plot-rp-source-targ-neg)), there appears to be a shift in the dynamics of these systems around observation 35, or May 3rd, 2012. To further investigate these local dynamics, we performed a sliding window technique known as windowed cross-recurrence quantification analysis (WCRQA).

## Windowed-Recurrence Quantification Analysis
### Syria as Target
The time period of differing dynamics, from observation 38 through 43, appears to be captured by WCRQA of social cohesion and count of negative events for both RR and DET (Figure \@ref(fig:plot-wcrqa-targ-all)). The amount of recurrence and determinism for these periods almost consistently lie above chance, suggesting that the intensity of negative events coincide with the degree of social cohesion found on Twitter, and that these states carry across multiple days in a way that is different from that of chance.
The above-chance levels of determinism and recurrence rate are demonstrating the coupling of social cohesion and counts of all and negative events during times of strife. We can see this string of highly violent events reflected in the mode ICEWS event for each day (Figure \@ref(fig:plot-mode-event)), where from May 1st, 2020 to May 15th, 2020, the most frequently occurring event was almost consistently -10, the most violent event in the coding system. These dates directly correlate with the above chance RR and DET metrics for the WCRQA of social cohesion and count of negative events. Furthermore, for WCRQA of social cohesion and positive event count during this time period, DET was entirely within chance (Figure \@ref(fig:plot-wcrqa-targ-all)). This supports our hypothesis, as the positive events are seemingly not coupled with the social cohesion on Twitter during civil uprisings during the Arab Spring.

### 2020-05-29 manuscript.tex
Figure and table cross-references

We first used cross-recurrence quantification analysis (CRQA; Zbilut et al., 1998) to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992), CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system (cf.~Nyquist frequency in signal processing; Grenander, 1959). We converted each time series (e.g., Figure 1) into deciles (e.g., Figure 2) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

@@

A recurrence plot (RP) is the visual representation of CRQA. To construct a recurrence plot, two time series are plotted against each other on the X and Y axes, and at every point where they share the same state, and point in placed on the plot to indicate recurrence. Visually analyzing these plots can help reveal patterns of structure and recurrence. The line y = x is known as the line of identity. This line falls where the time series match temporally, and a filled line would indicate perfect synchrony between the two time series. See Figures 3-5 for examples of RPs.

@@

To visualize WCRQA, the resulting metrics from each window can be plotted as a time series. This allows for visual interpretation of the pattern or fluctuation of the metrics across time. These dynamics can then be linked to other external events. See Figures 9-14 and 16-21 for examples.

@@

A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, maxL, and LAM. All other metrics were non-significant (Figure 3; Table 1). A permutation test for the CRQA of social cohesion and count of positive events found no statistical difference from chance for any of the metrics (Figure 4; Table 1). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, ENTR, and TT. All other metrics were non-significant (Figure 5; Table 1). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for target only data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endhead
DET & 28.525 & 0.000 & *** & 19.672 & 0.288 & & 25.902 & 0.001 & ***\tabularnewline
NRLINE & 80.000 & 0.000 & *** & 57.000 & 0.274 & & 71.000 & 0.005 & **\tabularnewline
maxL & 4.000 & 0.045 & * & 4.000 & 0.053 & . & 4.000 & 0.053 & .\tabularnewline
L & 2.175 & 0.082 & . & 2.105 & 0.474 & & 2.225 & 0.021 & *\tabularnewline
ENTR & 0.481 & 0.079 & . & 0.341 & 0.455 & & 0.575 & 0.015 & *\tabularnewline
rENTR & 0.438 & 0.352 & & 0.311 & 0.749 & & 0.523 & 0.135 &\tabularnewline
LAM & 35.902 & 0.002 & ** & 25.738 & 0.062 & . & 29.344 & 0.021 & *\tabularnewline
TT & 2.147 & 0.238 & & 2.492 & 0.012 & * & 2.295 & 0.049 & *\tabularnewline
\bottomrule
\end{longtable}

@@

A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, and LAM. All other metrics were non-significant (Figure 6; Table 2). A permutation test for the CRQA of social cohesion and count of positive events found a statistical difference from chance for DET, NRLINE, maxL, and LAM (Figure 7; Table 2). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, and ENTR. All other metrics were non-significant (Figure 8; Table 2). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{CRQA results for source and target data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endhead
DET & 25.246 & 0.001 & *** & 21.148 & 0.127 & & 22.623 & 0.023 & *\tabularnewline
NRLINE & 75.000 & 0.000 & *** & 59.000 & 0.193 & & 64.000 & 0.038 & *\tabularnewline
maxL & 3.000 & 0.423 & & 4.000 & 0.036 & * & 3.000 & 0.370 &\tabularnewline
L & 2.053 & 0.911 & & 2.186 & 0.060 & . & 2.156 & 0.155 &\tabularnewline
ENTR & 0.208 & 0.901 & & 0.510 & 0.047 & * & 0.433 & 0.165 &\tabularnewline
rENTR & 0.300 & 0.802 & & 0.464 & 0.255 & & 0.625 & 0.032 & *\tabularnewline
LAM & 34.262 & 0.003 & ** & 27.705 & 0.035 & * & 24.098 & 0.113 &\tabularnewline
TT & 2.247 & 0.107 & & 2.195 & 0.156 & & 2.100 & 0.360 &\tabularnewline
\bottomrule
\end{longtable}

@@

WCRQA was performed using the crqa package (Coco \& Dale, 2018) for R (R Core Team, 2019). Permutation tests were conducted for each WCRQA to establish upper and lower 95th and 99th percentiles. In the plots (see Figures 9-14 and 16-21), the 95th percentiles are represented by horizontal red lines and the 99th percentiles are represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there is a consistent period of above chance recurrence rate from the 31st through 48th window (Figure 9). Determinism is also almost consistently above chance from the 39th window onwards (Figure 10). For social cohesion and count of positive events there is a downward trend in RR as the window moves (Figure 11) and determinism stays almost entirely within chance across windows (Figure 12). For social cohesion and count of negative events, recurrence rate stay above or almost above chance from the 36th through the 48th window (Figure 13) and determinism is also above or nearly above chance from the 38th through 43rd window (Figure 14).

@@

For social cohesion and count of all events, there is a consistent period of above or nearly above chance determinism from the 26th through 46th window (Figure 16). Determinism also shows above or nearly above chance values in window spans 11-16 and 35-37 (Figure 17). For social cohesion and count of positive events, there is a downward trend in RR as the window moves (Figure 18) and determinism stays largely within chance with the exception of windows 13-19. For social cohesion and negative events, there appears to be an extended time period where RR (Figure 11) and DET (Figure 12) were greater than chance, from the 35th through 46th windows. These windows span from May 3rd, 2012 to May 14th, 2012.

@@

By adding Syria as a source in the data filtering, the dataset increased in size by 1,690 samples. These additional observations are times where Syria initiated an event with another country. Events initiated by Syria but directed internally within the country were already captured by the target only filtering. The increase in coupling between positive events and social cohesion is thus due to the inclusion of these externally directed events. Though CRQA metrics for count of all and count of negative events with social cohesion were largely unaffected, count of positive events and social cohesion showed significant levels of determinism, number of lines, max line length, and laminarity that were not present in the target filtered data. Given that the Syria during this time was still under Authoritarian rule, the government may have been initiating more positive external events to prevent intervention by other countries or try to cover up what was happening internally. We believe that the increased coupling of positive events is thus not reflective of actual social cohesion, but happened by chance. The inflation of the number of positive events when including source observations is causing increased coupling, but it is more likely that the Twitter users were using the platform to raise awareness of the violent, internal events that were being inflicted on them by their government officials and armies and not reporting on the positive external events that did not directly affect them or their safety. Further research should explore the semantic content of this tweet corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs; Figures 3-8), there appears to be a shift in the dynamics of these systems around observation 35, or May 3rd, 2012. To further investigate these local dynamics, we performed a sliding window technique known as windowed cross-recurrence quantification analysis (WCRQA).

@@

The time period of differing dynamics, from observation 38 through 43, appears to be captured by WCRQA of social cohesion and count of negative events for both RR (Figure 13) and DET (Figure 14). The amount of recurrence and determinism for these periods almost consistently lie above chance, suggesting that the intensity of negative events coincide with the degree of social cohesion found on Twitter, and that these states carry across multiple days in a way that is different from that of chance.
The above-chance levels of determinism and recurrence rate are demonstrating the coupling of social cohesion and counts of all and negative events during times of strife. We can see this string of highly violent events reflected in the mode ICEWS event for each day (Figure 15), where from May 1st, 2020 to May 15th, 2020, the most frequently occurring event was almost consistently -10, the most violent event in the coding system. These dates directly correlate with the above chance RR and DET metrics for the WCRQA of social cohesion and count of negative events. Furthermore, for WCRQA of social cohesion and positive event count during this time period, DET (Figure 12) was entirely within chance. This supports our hypothesis, as the positive events are seemingly not coupled with the social cohesion on Twitter during civil uprisings during the Arab Spring.

>>>>

We first used cross-recurrence quantification analysis (CRQA; Zbilut et al., 1998) to investigate the temporal patterns of social cohesion and real-world events in Syria during Arab Spring. As an extension of RQA, which captures the patterns of revisited states of a single time series (Webber Jr \& Zbilut, 1994; Zbilut \& Webber Jr, 1992), CRQA allows for two time series to be projected onto one another to investigate shared states and trajectories. Nonlinear analyses such as CRQA can reveal dynamics that are not captured by traditional linear methods. In order to perform CRQA, the time series need to be converted to the same scale, while still preserving a sampling rate that respects the native timescale of the system (cf.~Nyquist frequency in signal processing; Grenander, 1959). We converted each time series (e.g., Figure \ref{fig:raw-ts}) into deciles (e.g., Figure \ref{fig:deciled-ts}) to analyze the dynamics of different levels of social cohesion and counts of all, positive, and negative events on each day. In CRQA, a recurrent point is when two signals share the same state. These shared states can be at the same time or at later points in the time series.

@@

A recurrence plot (RP) is the visual representation of CRQA. To construct a recurrence plot, two time series are plotted against each other on the X and Y axes, and at every point where they share the same state, and point in placed on the plot to indicate recurrence. Visually analyzing these plots can help reveal patterns of structure and recurrence. The line y = x is known as the line of identity. This line falls where the time series match temporally, and a filled line would indicate perfect synchrony between the two time series. See Figures \ref{fig:plot-rp-targ-all}-\ref{fig:plot-rp-source-targ-neg} for examples of RPs.

@@

To visualize WCRQA, the resulting metrics from each window can be plotted as a time series. This allows for visual interpretation of the pattern or fluctuation of the metrics across time. These dynamics can then be linked to other external events. See Figures \ref{fig:plot-wcrqa-targ-all} and \ref{fig:plot-wcrqa-source-targ-all} for examples.

@@

A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, maxL, and LAM. All other metrics were non-significant (Figure \ref{fig:plot-rp-targ-all}; \autoref{table-1}). A permutation test for the CRQA of social cohesion and count of positive events found no statistical difference from chance for any of the metrics (Figure \ref{fig:plot-rp-targ-pos}; \autoref{table-1}). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, ENTR, and TT. All other metrics were non-significant (Figure \ref{fig:plot-rp-targ-neg}; \autoref{table-1}). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{\label{table-1}CRQA results for target only data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endhead
DET & 28.525 & 0.000 & *** & 19.672 & 0.288 & & 25.902 & 0.001 & ***\tabularnewline
NRLINE & 80.000 & 0.000 & *** & 57.000 & 0.274 & & 71.000 & 0.005 & **\tabularnewline
maxL & 4.000 & 0.045 & * & 4.000 & 0.053 & . & 4.000 & 0.053 & .\tabularnewline
L & 2.175 & 0.082 & . & 2.105 & 0.474 & & 2.225 & 0.021 & *\tabularnewline
ENTR & 0.481 & 0.079 & . & 0.341 & 0.455 & & 0.575 & 0.015 & *\tabularnewline
rENTR & 0.438 & 0.352 & & 0.311 & 0.749 & & 0.523 & 0.135 &\tabularnewline
LAM & 35.902 & 0.002 & ** & 25.738 & 0.062 & . & 29.344 & 0.021 & *\tabularnewline
TT & 2.147 & 0.238 & & 2.492 & 0.012 & * & 2.295 & 0.049 & *\tabularnewline
\bottomrule
\end{longtable}

@@

A permutation test for the CRQA of social cohesion and count of all events found a statistical difference from chance for DET, NRLINE, and LAM. All other metrics were non-significant (Figure \ref{fig:plot-rp-source-targ-all}; \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of positive events found a statistical difference from chance for DET, NRLINE, maxL, and LAM (Figure \ref{fig:plot-rp-source-targ-pos}; \autoref{table-2}). A permutation test for the CRQA of social cohesion and count of negative events found a statistical difference from chance for DET, NRLINE, maxL, L, and ENTR. All other metrics were non-significant (Figure \ref{fig:plot-rp-source-targ-neg}; \autoref{table-2}). Looking at the recurrence plots there appears, to be a shift in dynamics around the 36th observation which corresponds to May 4th, 2012.

\begin{longtable}[]{@{}cccccccccc@{}}
\caption{\label{table-2}CRQA results for source and target data.}\tabularnewline
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endfirsthead
\toprule
metric & all events & p & sig. & positive events & p & sig. & negative events & p & sig.\tabularnewline
\midrule
\endhead
DET & 25.246 & 0.001 & *** & 21.148 & 0.127 & & 22.623 & 0.023 & *\tabularnewline
NRLINE & 75.000 & 0.000 & *** & 59.000 & 0.193 & & 64.000 & 0.038 & *\tabularnewline
maxL & 3.000 & 0.423 & & 4.000 & 0.036 & * & 3.000 & 0.370 &\tabularnewline
L & 2.053 & 0.911 & & 2.186 & 0.060 & . & 2.156 & 0.155 &\tabularnewline
ENTR & 0.208 & 0.901 & & 0.510 & 0.047 & * & 0.433 & 0.165 &\tabularnewline
rENTR & 0.300 & 0.802 & & 0.464 & 0.255 & & 0.625 & 0.032 & *\tabularnewline
LAM & 34.262 & 0.003 & ** & 27.705 & 0.035 & * & 24.098 & 0.113 &\tabularnewline
TT & 2.247 & 0.107 & & 2.195 & 0.156 & & 2.100 & 0.360 &\tabularnewline
\bottomrule
\end{longtable}

@@

WCRQA was performed using the crqa package (Coco \& Dale, 2018) for R (R Core Team, 2019). Permutation tests were conducted for each WCRQA to establish upper and lower 95th and 99th percentiles. In the plots (see Figures \ref{fig:plot-wcrqa-targ-all} and \ref{fig:plot-wcrqa-source-targ-all}), the 95th percentiles are represented by horizontal red lines and the 99th percentiles are represented by horizontal orange lines. The trend lines are plotted in blue.

@@

For social cohesion and count of all events, there is a consistent period of above chance recurrence rate from the 31st through 48th window (Figure \ref{fig:plot-wcrqa-targ-all}). Determinism is also almost consistently above chance from the 39th window onwards (Figure \ref{fig:plot-wcrqa-targ-all}). For social cohesion and count of positive events there is a downward trend in RR as the window moves and determinism stays almost entirely within chance across windows (Figure \ref{fig:plot-wcrqa-targ-all}). For social cohesion and count of negative events, recurrence rate stay above or almost above chance from the 36th through the 48th window and determinism is also above or nearly above chance from the 38th through 43rd window (Figure \ref{fig:plot-wcrqa-targ-all}).

@@

For social cohesion and count of all events, there is a consistent period of above or nearly above chance determinism from the 26th through 46th window (Figure \ref{fig:plot-wcrqa-source-targ-all}). Determinism also shows above or nearly above chance values in window spans 11-16 and 35-37 (Figure \ref{fig:plot-wcrqa-source-targ-all}). For social cohesion and count of positive events, there is a downward trend in RR as the window moves and determinism stays largely within chance with the exception of windows 13-19 (Figure \ref{fig:plot-wcrqa-source-targ-all}). For social cohesion and negative events, there appears to be an extended time period where RR and DET were greater than chance, from the 35th through 46th windows (Figure \ref{fig:plot-wcrqa-source-targ-all}). These windows span from May 3rd, 2012 to May 14th, 2012.

@@

By adding Syria as a source in the data filtering, the dataset increased in size by 1,690 samples. These additional observations are times where Syria initiated an event with another country. Events initiated by Syria but directed internally within the country were already captured by the target only filtering. The increase in coupling between positive events and social cohesion is thus due to the inclusion of these externally directed events. Though CRQA metrics for count of all and count of negative events with social cohesion were largely unaffected, count of positive events and social cohesion showed significant levels of determinism, number of lines, max line length, and laminarity that were not present in the target filtered data. Given that the Syria during this time was still under Authoritarian rule, the government may have been initiating more positive external events to prevent intervention by other countries or try to cover up what was happening internally. We believe that the increased coupling of positive events is thus not reflective of actual social cohesion, but happened by chance. The inflation of the number of positive events when including source observations is causing increased coupling, but it is more likely that the Twitter users were using the platform to raise awareness of the violent, internal events that were being inflicted on them by their government officials and armies and not reporting on the positive external events that did not directly affect them or their safety. Further research should explore the semantic content of this tweet corpus to determine what the socially cohered language was focused on.
By visual inspection of the recurrence plots (RPs; Figures \ref{fig:plot-rp-targ-all}-\ref{fig:plot-rp-source-targ-neg}), there appears to be a shift in the dynamics of these systems around observation 35, or May 3rd, 2012. To further investigate these local dynamics, we performed a sliding window technique known as windowed cross-recurrence quantification analysis (WCRQA).

@@

The time period of differing dynamics, from observation 38 through 43, appears to be captured by WCRQA of social cohesion and count of negative events for both RR and DET (Figure \ref{fig:plot-wcrqa-targ-all}). The amount of recurrence and determinism for these periods almost consistently lie above chance, suggesting that the intensity of negative events coincide with the degree of social cohesion found on Twitter, and that these states carry across multiple days in a way that is different from that of chance.
The above-chance levels of determinism and recurrence rate are demonstrating the coupling of social cohesion and counts of all and negative events during times of strife. We can see this string of highly violent events reflected in the mode ICEWS event for each day (Figure \ref{fig:plot-mode-event}), where from May 1st, 2020 to May 15th, 2020, the most frequently occurring event was almost consistently -10, the most violent event in the coding system. These dates directly correlate with the above chance RR and DET metrics for the WCRQA of social cohesion and count of negative events. Furthermore, for WCRQA of social cohesion and positive event count during this time period, DET was entirely within chance (Figure \ref{fig:plot-wcrqa-targ-all}). This supports our hypothesis, as the positive events are seemingly not coupled with the social cohesion on Twitter during civil uprisings during the Arab Spring.
